# Child poverty

The Child Poverty (Scotland) Act 2017 requires us to report annually on a number of child poverty measures. The latest single-year estimates can be found in the [Child Poverty Update](cpupdate.html). The analysis in this report however includes three-year averaged estimates only, which are more suitable for showing trends.

`r infobox("In this publication, 'child' refers to a dependent child. This is explained in the [Definitions section](index.html#child).")`

## Key trends {.tabset}

### Relative poverty

**Relative poverty rate for children no longer rising**

It is estimated that 24% of children (230,000 children each year) were living in relative poverty after housing costs in 2016-19. Before housing costs, it is estimated that 20% of children (200,000 children each year) were in relative poverty.

After a long fall between 1999-02 and 2010-13, which halted briefly just before the recession, child poverty rates had been rising again but did not rise in the latest periods.

#### Figure 10 - Proportion of children in relative poverty, Scotland
```{r chart10}
interactive(povertycharts[["chart10"]])
```

### Absolute poverty

**Absolute poverty rate for children stable**

Absolute child poverty after housing costs affected 21% (210,000 children each year). Before housing costs, absolute child poverty was at 17% (160,000 children each year).

Absolute child poverty before housing costs steadily decreased between the mid-nineties and just before the recession. The decrease of the after housing costs rate then slowed down. The before housing costs measure continued to decrease until 2013-16 and has been largely flat since. As a result, the gap between absolute child poverty before and after housing costs widened in recent years.

#### Figure 11 - Proportion of children in absolute poverty, Scotland
```{r chart11}
interactive(povertycharts[["chart11"]], height = 4)
```

### Material deprivation

**Children in combined material deprivation and low income broadly stable**

It is estimated that 12% of children (120,000 children each year) were living in combined low income and material deprivation after housing costs in 2016-19. Before housing costs, this was 11% of children (110,000 children).

Combined low income and child material deprivation is an additional way of measuring living standards and refers to the inability of households to afford basic goods and activities that are seen as necessities in society.

More detail on this can be found in the [Definitions section](index.html#matdep).

#### Figure 12 - Proportion of children in combined low income and material deprivation, Scotland
```{r chart12}
interactive(povertycharts[["chart12"]])
```

`r infobox("Material deprivation data has been collected since 2004/05. Due to a change in the methodology, it is not possible to compare the most recent years with years before 2010-13. The break in the time series indicates this change.")`

### In-work poverty

**In-work poverty for children no longer rising**

While the poverty risk is much lower for children in working households compared to those in non-working households, not all work pays enough to lift the household above the poverty threshold.

It is estimated that in 2016-19, 65% of children in relative poverty after housing costs (150,000 children each year) were living in working households. Before housing costs, 64% of children in poverty (130,000 children) were living in working households.

This data suggests that after a steady and steep increase of in-work poverty since 2007-10, in-work poverty is now stable at a high level.

`r infobox("The terms 'working' and 'in-work poverty' here refer to paid employment only. In-work poverty refers to people living in households where at least one member of the household is in either full or part-time paid work, but where the household income is below the relative poverty threshold.")`

#### Figure 13 - Proportion of children in relative poverty who live in working households, Scotland
```{r chart13}
interactive(povertycharts[["chart13"]])
```

### Food insecurity

Add commentary here: what is food insecurity, previous data collection, differences to UK measure, reference period, etc.

#### Figure 14 - Proportion of children with food insecurity, Scotland 2018/19
```{r chart14}
#interactive(povertycharts[["chart14"]])
```

Bar chart showing the proportion of children who've expericenced food insecurity, broken down by poverty status (all children, childrenin relative poverty AHC, children in combined low income and material deprivation)

## Priority groups {.tabset}

add commentary here explaining where the groups are coming from, links with TCPDP, etc.

### Relative poverty

#### Figure 15 - Proportion of children in relative poverty after housing cost, Scotland
```{r chart15}
interactive(povertycharts[["chart15"]])
```

### Absolute poverty

#### Figure 15b - Proportion of children in absolute poverty after housing costs, Scotland
```{r chart15b}
interactive(povertycharts[["chart15b"]])
```

### Material deprivation

#### Figure 15c - Proportion of children in combined low income and material deprivation, Scotland
```{r chart15c}
interactive(povertycharts[["chart15c"]])
```

