## Income inequality {.tabset}

### Palma

```{r}
keymessage <- "Palma ratio of income inequality fluctuating"
message(keymessage)

rateB <- filter(data$palma, yearn == max(yearn), key == "Before housing costs") %>%
  mutate(y = fmtpct(round2(y - 1, 2))) %>% select(y) %>% pull()
```

The Palma ratio of income inequality is the total income of the top ten percent of the population divided by the total income of the bottom forty percent of the population (written as a percentage). It is commonly used to estimate how much more income top-income households have compared to those at the bottom. 

The Palma ratio is usually calculated from income before housing costs, but we have included it for after-housing-costs income as well. After-housing-costs incomes are distributed more unequally.

In `r params$data_period`, the top ten percent of the population had `r rateB` more income (before housing costs) than the bottom forty percent combined. This compares to 33% more income of the top ten percent in 2007-10, the period with the highest income inequality in this time series as measured by the Palma ratio.

Income inequality has been fluctuating since the beginning of this data collection in the mid-nineties, and, after the most recent high point in 2015-18 is currently decreasing again.

```{r chart27}
title <- str_c("Figure 27: ", keymessage)
subtitle <- "Palma ratio of income inequality, Scotland"
interactive(povertycharts$chart27, title = title, subtitle = subtitle)
```

Income decile shares are also available in the associated tables.

### Gini

```{r}
keymessage <- "Gini coefficient of income inequality fluctuating"
message(keymessage)

rateB <- filter(data$gini, yearn == max(yearn), key == "Before housing costs") %>%
   select(label) %>% pull()
```

The Gini coefficient measures income inequality on a scale from 0% to 100%. A Gini of 100% means that only one person has an income, and everyone else has none. A Gini of 0% means that everyone has the same income. 

In `r params$data_period`, the Gini coefficient (before housing costs) was `r rateB`. The Gini coefficient is usually calculated from income before housing costs, but we have included it for after housing costs income as well.  After-housing-costs incomes are distributed more unequally.

Income inequality has been fluctuating since the beginning of this data collection in the mid-nineties, and, after the most recent high point in 2015-18 is currently decreasing again.

```{r chart28}
title <- str_c("Figure 28: ", keymessage)
subtitle <- "Gini coefficient of income inequality, Scotland"
interactive(povertycharts$chart28, title = title, subtitle = subtitle)
```
