## Working-age adults

Working-age adults tend to be less likely to be in poverty compared to children.

Working-age adults are all adults up to the state pension age. Women’s state pension age reached 65 in November 2018, aligning it with men’s state pension age. Since December 2018, the state pension age for both men and women has been increasing. In the latest data period included in this report, the State Pension age for both men and women increased to 66 years.

### Relative poverty {#working-age-relative-poverty}

```{r wa_rel}

keymessage <- "Relative poverty rate for working-age adults largely stable"
fign <- fign + 1
title <- paste0("Figure ", fign, ": ", keymessage)
subtitle <- "Proportion of working-age adults in relative poverty, Scotland<br>Note: see commentary for data concerns around latest estimate"

chart <- povertycharts$wa_rel %>% 
  highcharter::hc_title(text = title) %>% 
  highcharter::hc_subtitle(text = subtitle) %>% 
  highcharter::hc_exporting(filename = paste0("fig_", fign)) %>% 
  highcharter::hc_credits(enabled = TRUE, text = source,
             href = href)

table1 <- datatable(data = htmltables$wa_rel,
                            id = paste0("table-", fign, "a"),
                            caption = paste0("Table ", fign, "a: ", subtitle),
                            source = source,
                            options = list(rowHeaders = TRUE))

table2 <- datatable(data = htmltables$wa_samples,
                            id = paste0("table-", fign, "b"),
                            caption = paste0("Table ", fign, "b: Sample size"),
                            source = source,
                            options = list(rowHeaders = TRUE))
rateA <- filter(htmltables$wa_rel, Measure == "After housing costs")[[params$data_period]]
numberA <- scales::comma(filter(povertytables$relAHC$numbers, Group == "Working-age adults")[[params$data_period]])
rateB <- filter(htmltables$wa_rel, Measure == "Before housing costs")[[params$data_period]]
numberB <- scales::comma(filter(povertytables$relBHC$numbers, Group == "Working-age adults")[[params$data_period]])

description <- paste0("Relative poverty for working-age adults has been broadly stable since the nineties, when reporting began. The relative poverty rate in ", params$data_period, " was estimated to be ", rateA, " after housing costs, and ", rateB, " before housing costs.

The latest estimate is higher than in previous years, driven by more working-age adults who are economically inactive. This means they are not in work and are not looking for work, maybe because they are disabled or have retired early. It is yet unclear whether this is the beginning of a new trend, or if it is a volatile data point. It might be the latter, because UK data shows declining working-age poverty, and Scottish working-age income trends usually move similarly to the UK's.

In ", params$data_period, ", there were ", numberA, " working-age adults in poverty after housing costs, and ", numberB, " before housing costs.",
infobox("Add confidence intervals by clicking on the greyed-out parts of the legend."))

```

```{r wa_rel_chart}

message(keymessage)
htmltools::tags$figure(id = paste0("figure-", fign),
                       htmltools::tags$div(class = "highcharts-description",
                                           id = paste0("figure-", fign, "-description"),
                                           create_html(description)),
                       chart)
tables_panel(list(table1, table2))

```

#### Working poverty

Having paid work is an effective way out of poverty, and those families where all adults are in full-time work have a low poverty risk. But having a job is not always enough, for example when it does not pay well, or when someone is unable to work enough hours.

```{r wa_work}

keymessage <- "Most working-age adults in poverty live in working households"

fign <- fign + 1

title <- paste0("Figure ", fign, "a: ", keymessage)
title2 <- paste0("Figure ", fign, "b: ", keymessage)
subtitle <- "Working-age adults in relative poverty after housing costs by household work status, Scotland"
subtitle2 <- "Share of working-age adults in relative poverty after housing costs who live in working households, Scotland<br>Note: see commentary for data concerns around latest estimate"

chart <- povertycharts$wa_work %>% 
  highcharter::hc_title(text = title) %>% 
  highcharter::hc_subtitle(text = subtitle) %>% 
  highcharter::hc_exporting(filename = paste0("fig_", fign, "a")) %>% 
  highcharter::hc_credits(enabled = TRUE, text = source,
             href = href)

chart2 <- povertycharts$wa_work2 %>% 
  highcharter::hc_title(text = title2) %>% 
  highcharter::hc_subtitle(text = subtitle2) %>% 
  highcharter::hc_exporting(filename = paste0("fig_", fign, "b")) %>% 
  highcharter::hc_credits(enabled = TRUE, text = source,
             href = href)

table1 <- datatable(data = htmltables$wa_work,
                            id = paste0("table-", fign, "a"),
                            caption = paste0("Table ", fign, "a: ", subtitle),
                            source = source,
                            options = list(rowHeaders = TRUE))

table2 <- datatable(data = htmltables$wa_work_samples,
                            id = paste0("table-", fign, "b"),
                            caption = paste0("Table ", fign, "b: Sample size"),
                            source = source,
                            options = list(rowHeaders = TRUE))

rateA <- filter(htmltables$wa_work, Group == "In working households")[[params$data_period]]
numberA <- scales::comma(filter(povertytables$workinghh_wa$rel_numbers, Group == "Someone in paid work")[[params$data_period]])

description <- paste0("In ", params$data_period, ", ", rateA, " of working-age adults (", numberA, " working-age adults each year) in relative poverty after housing costs were living in a household where someone was in paid work.

The share of working-age adults in poverty who lived in working households continuously increased since 2011-14, reflecting high employment rates, and now appears largely stable except for the latest estimate. 

While the number of adults in working poverty remains largely stable, their share of all working-age adults in poverty shows a drop. This is because the additional adults who are now in poverty largely live in non-working households. It is yet unclear whether this is the beginning of a new trend, or if it is a volatile data point.

", infobox("The terms ‘working’ and ‘in-work poverty’ here refer to paid employment. They do not include unpaid work such as caring for your children or other family members. In-work poverty refers to the share of children in poverty who live in households where at least one member of the household is in either full or part-time paid work."))

```

```{r wa_work_chart}

message(keymessage)
htmltools::tags$figure(id = paste0("figure-", fign),
                       htmltools::tags$div(class = "highcharts-description",
                                           id = paste0("figure-", fign, "-description"),
                                           create_html(description)),
                       chart, chart2)
tables_panel(list(table1, table2))

```

### Absolute poverty {#working-age-absolute-poverty}

```{r wa_abs}

keymessage <- "Absolute poverty rate for working-age adults stable"

fign <- fign + 1

title <- paste0("Figure ", fign, ": ", keymessage)
subtitle <- "Proportion of working-age adults in absolute poverty, Scotland<br>Note: see commentary for data concerns around latest estimate"

chart <- povertycharts$wa_abs %>% 
  highcharter::hc_title(text = title) %>% 
  highcharter::hc_subtitle(text = subtitle) %>% 
  highcharter::hc_exporting(filename = paste0("fig_", fign)) %>% 
  highcharter::hc_credits(enabled = TRUE, text = source,
             href = href)

table1 <- datatable(data = htmltables$wa_abs,
                            id = paste0("table-", fign, "a"),
                            caption = paste0("Table ", fign, "a: ", subtitle),
                            source = source,
                            options = list(rowHeaders = TRUE))

table2 <- datatable(data = htmltables$wa_samples,
                            id = paste0("table-", fign, "b"),
                            caption = paste0("Table ", fign, "b: Sample size"),
                            source = source,
                            options = list(rowHeaders = TRUE))

rateA <- filter(htmltables$wa_abs, Measure == "After housing costs")[[params$data_period]]
numberA <- scales::comma(filter(povertytables$absAHC$numbers, Group == "Working-age adults")[[params$data_period]])
rateB <- filter(htmltables$wa_abs, Measure == "Before housing costs")[[params$data_period]]
numberB <- scales::comma(filter(povertytables$absBHC$numbers, Group == "Working-age adults")[[params$data_period]])

description <- paste0("Absolute poverty amongst working-age adults remained broadly stable during the last fifteen years. In ", params$data_period, ", ", rateA, " of working-age adults were in absolute poverty after housing costs, and ", rateB, " before housing costs.

This means that in ", params$data_period, ", there were ", numberA, " working-age adults each year in absolute poverty after housing costs, compared to ", numberB, " before housing costs.

The latest estimates show an increase, but again, it is yet unclear whether this is the beginning of a new trend, or if it is a volatile data point.")

```

```{r wa_abs_chart}

message(keymessage)
htmltools::tags$figure(id = paste0("figure-", fign),
                       htmltools::tags$div(class = "highcharts-description",
                                           id = paste0("figure-", fign, "-description"),
                                           create_html(description)),
                       chart)
tables_panel(list(table1, table2))

```

### Food security {#working-age-food-security}

```{r wa_foodsec}

keymessage <- "More than one third of working-age adults in poverty live in households lacking high food security"

fign <- fign + 1

title <- paste0("Figure ", fign, ": ", keymessage)
subtitle <- str_c("Levels of food security of all working-age adults and those in poverty after housing costs, Scotland ",
                 params$data_period)

chart <- povertycharts$wa_foodsec %>% 
  highcharter::hc_title(text = title) %>% 
  highcharter::hc_subtitle(text = subtitle) %>% 
  highcharter::hc_exporting(filename = paste0("fig_", fign)) %>% 
  highcharter::hc_credits(enabled = TRUE, text = source,
             href = href)

table1 <- datatable(data = htmltables$wa_foodsec,
                            id = paste0("table-", fign, "a"),
                            caption = paste0("Table ", fign, "a: ", subtitle),
                            source = source,
                            options = list(rowHeaders = TRUE))

table2 <- datatable(data = htmltables$wa_foodsec_samples,
                            id = paste0("table-", fign, "b"),
                            caption = paste0("Table ", fign, "b: Sample size"),
                            source = source,
                            options = list(rowHeaders = TRUE))

rateH_all <- filter(htmltables$wa_foodsec, Group == "All working-age adults")[["High"]]
rateL_all <- fmtpct(1 - filter(povertytables$foodsec$comps, Group == "Working-age adults")[["High"]])
rateH_rel <- filter(htmltables$wa_foodsec, Group == "In relative poverty")[["High"]]
rateH_sev <- filter(htmltables$wa_foodsec, Group == "In severe poverty")[["High"]]

description <- paste0("Food security is measured at a household level. The person in the household who knows most about buying and preparing food responds to the questions about food security. Note that not everyone in the household may experience this in the same way. For example, a parent may have worried about running out of food or reduced their own meal sizes, but protected their young children from this experience. 

In ", params$data_period, ", ", rateH_all, " of working-age adults lived in households with high food security. This means that ", rateL_all, " of working-age adults lived in households with marginal, low or very low food security.

Working-age adults in poverty were less likely to have high food security: just ", rateH_rel, " of those in relative poverty, and ", rateH_sev, " of those in [severe poverty](#Relative_absolute_and_severe_poverty) lived in high food security households. 

People are in severe poverty when their household income is less than half of the UK median income.

", infobox("Household food security questions were newly added to the Family Resources Survey in 2019/20. They ask about whether people were worried about running out of food, had to reduce meal sizes or skip meals. More information can be found in the [Definitions section](#food-security-definition)."))

```

```{r wa_foodsec_chart}

message(keymessage)
htmltools::tags$figure(id = paste0("figure-", fign),
                       htmltools::tags$div(class = "highcharts-description",
                                           id = paste0("figure-", fign, "-description"),
                                           create_html(description)),
                       chart)
tables_panel(list(table1, table2))

```

