## Equality analysis

Poverty is measured at a household level. Everyone in the same household is considered either in poverty or not in poverty. This makes it difficult to measure the poverty risk by individual characteristics such as age or gender for people who share the households with others. 

For most characteristics in this section, we include everyone in the analysis, but keep in mind that the poverty risk is influenced by others in the household.

`r infobox("Note that estimates in this section tend to fluctuate, because some groups are smaller and have small sample sizes. This means that we can comment on consistent differences between groups, but any short-term changes over time are hidden in the fluctuations.")`

### Age

The age analysis looks at the age of the [head of the household](#household-head). Poverty rates refer to the proportion of people in households by age of the household head.

#### Working-age households

```{r wa_age}

keymessage <- "Relative poverty rates highest for the youngest households"

fign <- fign + 1

title <- paste0("Figure ", fign, ": ", keymessage)
subtitle <- "Proportion of people in relative poverty after housing costs, Scotland"

chart <- povertycharts$wa_age %>% 
  highcharter::hc_title(text = title) %>% 
  highcharter::hc_subtitle(text = subtitle) %>% 
  highcharter::hc_exporting(filename = paste0("fig_", fign)) %>% 
  highcharter::hc_credits(enabled = TRUE, text = source, href = href)

table1 <- datatable(data = htmltables$wa_age,
                            id = paste0("table-", fign, "a"),
                            caption = paste0("Table ", fign, "a: ", subtitle),
                            source = source,
                            options = list(rowHeaders = TRUE))

table2 <- datatable(data = htmltables$wa_age_samples,
                            id = paste0("table-", fign, "b"),
                            caption = paste0("Table ", fign, "b: Sample size"),
                            source = source,
                            options = list(rowHeaders = TRUE))

rate_16_24 <- filter(htmltables$wa_age, Group == "16-24")[[params$data_period]]
rate_25_34 <- filter(htmltables$wa_age, Group == "25-34")[[params$data_period]]
rate_65 <- filter(htmltables$wa_age, Group == "65+")[[params$data_period]]

num_16_24 <- scales::comma(filter(povertytables$ageband_pp$rel_numbers, Group == "16-24")[[params$data_period]])
num_25_34 <- scales::comma(filter(povertytables$ageband_pp$rel_numbers, Group == "25-34")[[params$data_period]])
num_65 <- scales::comma(filter(povertytables$ageband_pp$rel_numbers, Group == "65+")[[params$data_period]])

description <- paste0("In the last 15 years, the youngest households (household heads aged 16-24) have been consistently more likely to be in relative poverty compared to older households.

Figure ", fign, " shows that in ", params$data_period, ", ", rate_16_24, " of people in households with household heads aged 16-24 were in relative poverty after housing costs (", num_16_24, " people each year).

In comparison, the age groups 25-34, 35-44, 45-54 and 55-64 all had similar (and lower) poverty rates between 17% and 22%.")

```

```{r wa_age_chart}

message(keymessage)
htmltools::tags$figure(id = paste0("figure-", fign),
                       htmltools::tags$div(class = "highcharts-description",
                                           id = paste0("figure-", fign, "-description"),
                                           create_html(description)),
                       chart)
tables_panel(list(table1, table2))

```

#### Pensioner households

```{r pn_age}

fign <- fign + 1

keymessage <- "Poverty rates similar among older age groups"

title <- paste0("Figure ", fign, ": ", keymessage)
subtitle <- "Proportion of people in relative poverty after housing costs, Scotland"

chart <- povertycharts$pn_age %>% 
  highcharter::hc_title(text = title) %>% 
  highcharter::hc_subtitle(text = subtitle) %>% 
  highcharter::hc_exporting(filename = paste0("fig_", fign)) %>% 
  highcharter::hc_credits(enabled = TRUE, text = source, href = href)

table1 <- datatable(data = htmltables$pn_age,
                            id = paste0("table-", fign, "a"),
                            caption = paste0("Table ", fign, "a: ", subtitle),
                            source = source,
                            options = list(rowHeaders = TRUE))

table2 <- datatable(data = htmltables$pn_age_samples,
                            id = paste0("table-", fign, "b"),
                            caption = paste0("Table ", fign, "b: Sample size"),
                            source = source,
                            options = list(rowHeaders = TRUE))

description <- paste0("Older households tend to have lower poverty rates compared to working-age households. There were no marked differences between households with heads aged 65-74, 75-84, and 85 and older.")

```

```{r pn_age_chart}

message(keymessage)
htmltools::tags$figure(id = paste0("figure-", fign),
                       htmltools::tags$div(class = "highcharts-description",
                                           id = paste0("figure-", fign, "-description"),
                                           create_html(description)),
                       chart)
tables_panel(list(table1, table2))

```

Poverty estimates for children by age are available in the associated tables.

### Gender

```{r gender}

rate_all <- fmtpct(filter(povertytables$marital_ad$rel_rates, Group == "All")[[params$data_period]])
rate_single <- fmtpct(filter(povertytables$singlehh_ad$rel_rates, Group == "All")[[params$data_period]])

rate_allmen <- fmtpct(filter(povertytables$sex_ad$rel_rates, Group == "Male")[[params$data_period]])

rate_allwomen <- fmtpct(filter(povertytables$sex_ad$rel_rates, Group == "Female")[[params$data_period]])

```

In `r params$data_period`, `r rate_allwomen` of women and `r rate_allmen` of men were in relative poverty after housing costs. 

Poverty is measured at a household level. This means that men and women in the same household are both either in poverty or not in poverty. In the analysis below, we therefore only include single adult households (with or without dependent children).

In `r params$data_period`, the relative poverty rate after housing costs for single adults (working-age and pensioners) was `r rate_single`, higher than for the total adult population (`r rate_all`).

#### Single working-age adults and gender

```{r wa_gender}

keymessage <- "Relative poverty rates highest for single mothers and single men"

fign <- fign + 1

title <- paste0("Figure ", fign, ": ", keymessage)
subtitle <- "Proportion of single working-age adults in relative poverty after housing costs, Scotland"

chart <- povertycharts$wa_gender %>% 
  highcharter::hc_title(text = title) %>% 
  highcharter::hc_subtitle(text = subtitle) %>% 
  highcharter::hc_exporting(filename = paste0("fig_", fign)) %>% 
  highcharter::hc_credits(enabled = TRUE, text = source, href = href)

table1 <- datatable(data = htmltables$wa_gender,
                            id = paste0("table-", fign, "a"),
                            caption = paste0("Table ", fign, "a: ", subtitle),
                            source = source,
                            options = list(rowHeaders = TRUE))

table2 <- datatable(data = htmltables$wa_gender_samples,
                            id = paste0("table-", fign, "b"),
                            caption = paste0("Table ", fign, "b: Sample size"),
                            source = source,
                            options = list(rowHeaders = TRUE))

rate_singlemother <- fmtpct(filter(povertytables$singlehh_ad$rel_rates, Group == "Female working-age adult with dependent children")[[params$data_period]])
rate_singlewoman <- fmtpct(filter(povertytables$singlehh_ad$rel_rates, Group == "Female working-age adult, no dependent children")[[params$data_period]])
rate_singleman <- fmtpct(filter(povertytables$singlehh_ad$rel_rates, Group == "Male working-age adult, no dependent children")[[params$data_period]])

num_singlemother <- scales::comma(filter(povertytables$singlehh_ad$rel_numbers, Group == "Female working-age adult with dependent children")[[params$data_period]])
num_singlewoman <- scales::comma(filter(povertytables$singlehh_ad$rel_numbers, Group == "Female working-age adult, no dependent children")[[params$data_period]])
num_singleman <- scales::comma(filter(povertytables$singlehh_ad$rel_numbers, Group == "Male working-age adult, no dependent children")[[params$data_period]])

description <- paste0("The poverty rate was highest for single mothers (", rate_singlemother, ", number not available) and single childless men (", rate_singleman, ", ", num_singleman, "). The poverty rate for single childless women was ", rate_singlewoman, " (", num_singlewoman, "). Estimates for single fathers are not available due to small sample sizes.

", infobox("In this publication, 'child' refers to a dependent child living in the household. This is explained in the [Definitions section](#dependent-children)."))

```

```{r wa_gender_chart}

message(keymessage)
htmltools::tags$figure(id = paste0("figure-", fign),
                       htmltools::tags$div(class = "highcharts-description",
                                           id = paste0("figure-", fign, "-description"),
                                           create_html(description)),
                       chart)
tables_panel(list(table1, table2))

```

#### Single pensioners and gender

```{r pn_gender}

keymessage <- "Relative poverty rates similar for male and female pensioners"

fign <- fign + 1

title <- paste0("Figure ", fign, ": ", keymessage)
subtitle <- "Proportion of single pensioners in relative poverty after housing costs, Scotland"

chart <- povertycharts$pn_gender %>% 
  highcharter::hc_title(text = title) %>% 
  highcharter::hc_subtitle(text = subtitle) %>% 
  highcharter::hc_exporting(filename = paste0("fig_", fign)) %>% 
  highcharter::hc_credits(enabled = TRUE, text = source, href = href)

table1 <- datatable(data = htmltables$pn_gender,
                            id = paste0("table-", fign, "a"),
                            caption = paste0("Table ", fign, "a: ", subtitle),
                            source = source,
                            options = list(rowHeaders = TRUE))

table2 <- datatable(data = htmltables$pn_gender_samples,
                            id = paste0("table-", fign, "b"),
                            caption = paste0("Table ", fign, "b: Sample size"),
                            source = source,
                            options = list(rowHeaders = TRUE))

rate_fem <- fmtpct(filter(povertytables$singlehh_ad$rel_rates, Group == "Female pensioner")[[params$data_period]])
rate_man <- fmtpct(filter(povertytables$singlehh_ad$rel_rates, Group == "Male pensioner")[[params$data_period]])

num_fem <- scales::comma(filter(povertytables$singlehh_ad$rel_numbers, Group == "Female pensioner")[[params$data_period]])

rate_men80 <- filter(povertytables$median_ages$gender, singlehh == "Male pensioner")[["share"]]
rate_women80 <- filter(povertytables$median_ages$gender, singlehh == "Female pensioner")[["share"]]

description <- paste0("In ", params$data_period, ", ", rate_fem, " of single female pensioners (", num_fem, " women each year) and ", rate_man, " of single male pensioners (number not available) were in relative poverty after housing costs. Note that there were too few single male pensioners in poverty in the sample to produce a robust estimate of their number.

In most years, the poverty rate after housing costs for single female pensioners had been higher than that for single male pensioners.")

```

```{r pn_gender_chart}

message(keymessage)
htmltools::tags$figure(id = paste0("figure-", fign),
                       htmltools::tags$div(class = "highcharts-description",
                                           id = paste0("figure-", fign, "-description"),
                                           create_html(description)),
                       chart)
tables_panel(list(table1, table2))

```

Estimates for all men and women (whether in single, couple or other households) are available in the associated tables.

### Sexual orientation

```{r sexid}

fign <- fign + 1

keymessage <- "Relative poverty rates higher for LGB+ adults"

title <- paste0("Figure ", fign, ": ", keymessage)
subtitle <- "Proportion of adults in relative poverty after housing costs, Scotland"

chart <- povertycharts$ad_sexid %>% 
  highcharter::hc_title(text = title) %>% 
  highcharter::hc_subtitle(text = subtitle) %>% 
  highcharter::hc_exporting(filename = paste0("fig_", fign)) %>% 
  highcharter::hc_credits(enabled = TRUE, text = source, href = href)

table1 <- datatable(data = htmltables$ad_sexid,
                            id = paste0("table-", fign, "a"),
                            caption = paste0("Table ", fign, "a: ", subtitle),
                            source = source,
                            options = list(rowHeaders = TRUE))

table2 <- datatable(data = htmltables$ad_sexid_age,
                            id = paste0("table-", fign, "b"),
                            caption = paste0("Table ", fign, "b: Median age"),
                            source = source,
                            options = list(rowHeaders = TRUE))

table3 <- datatable(data = htmltables$ad_sexid_samples,
                            id = paste0("table-", fign, "c"),
                            caption = paste0("Table ", fign, "c: Sample size"),
                            source = source,
                            options = list(rowHeaders = TRUE))

rate_lgb <- fmtpct(filter(povertytables$sexid_ad$rel_rates, Group == "Other")[[params$data_period]])
rate_straight <- fmtpct(filter(povertytables$sexid_ad$rel_rates, Group == "Heterosexual / straight")[[params$data_period]])
rate_miss <- fmtpct(filter(povertytables$sexid_ad$rel_rates, Group == "(Missing)")[[params$data_period]])

num_lgb <- scales::comma(filter(povertytables$sexid_ad$rel_numbers, Group == "Other")[[params$data_period]])
num_straight <- scales::comma(filter(povertytables$sexid_ad$rel_numbers, Group == "Heterosexual / straight")[[params$data_period]])
num_miss <- scales::comma(filter(povertytables$sexid_ad$rel_numbers, Group == "(Missing)")[[params$data_period]])

description <- paste0("The poverty rate has been consistently higher for LGB+ adults compared to straight / heterosexual adults. In ", params$data_period, ", ", rate_lgb, " of LGB+ adults (number not available) were in poverty, compared to ", rate_straight, " of straight adults (", num_straight, " adults) and ", rate_miss, " of adults whose sexual orientation we don't know (", num_miss, " adults).

We included the 'Missing' category in the tables because it contains a fairly large number of adults. This group contains adults who were not asked the question because they weren't present at the interview. It also includes adults who chose not to answer the question.  

Note that there were too few LGB+ adults in poverty in the sample to produce a robust estimate of their population. Also, measurement uncertainty is quite wide for this group.

", infobox("LGB+ adults in this analysis are those adults who responded that they thought of themselves as gay, lesbian, bisexual or something other than straight or heterosexual."))

```

```{r sexid_chart}

message(keymessage)
htmltools::tags$figure(id = paste0("figure-", fign),
                       htmltools::tags$div(class = "highcharts-description",
                                           id = paste0("figure-", fign, "-description"),
                                           create_html(description)),
                       chart)
tables_panel(list(table1, table2))

```

### Marital status

```{r marital}
keymessage <- "Relative poverty rates highest for single, divorced & separated, and lowest for married adults"

fign <- fign + 1

title <- paste0("Figure ", fign, ": ", keymessage)
subtitle <- "Proportion of adults in relative poverty after housing costs, Scotland"

chart <- povertycharts$ad_marital %>% 
  highcharter::hc_title(text = title) %>% 
  highcharter::hc_subtitle(text = subtitle) %>% 
  highcharter::hc_exporting(filename = paste0("fig_", fign)) %>% 
  highcharter::hc_credits(enabled = TRUE, text = source, href = href)

table1 <- datatable(data = htmltables$ad_marital,
                            id = paste0("table-", fign, "a"),
                            caption = paste0("Table ", fign, "a: ", subtitle),
                            source = source,
                            options = list(rowHeaders = TRUE))

table2 <- datatable(data = htmltables$ad_marital_samples,
                            id = paste0("table-", fign, "b"),
                            caption = paste0("Table ", fign, "b: Sample size"),
                            source = source,
                            options = list(rowHeaders = TRUE))

rate_single <- fmtpct(filter(povertytables$marital_ad$rel_rates, Group == "Single")[[params$data_period]])
rate_divorced <- fmtpct(filter(povertytables$marital_ad$rel_rates, Group == "Divorced / Civil Partnership dissolved / separated")[[params$data_period]])
rate_married <- fmtpct(filter(povertytables$marital_ad$rel_rates, Group == "Married / Civil Partnership")[[params$data_period]])
rate_widowed <- fmtpct(filter(povertytables$marital_ad$rel_rates, Group == "Widowed")[[params$data_period]])
rate_cohab <- fmtpct(filter(povertytables$marital_ad$rel_rates, Group == "Cohabiting")[[params$data_period]])

num_single <- scales::comma(filter(povertytables$marital_ad$rel_numbers, Group == "Single")[[params$data_period]])
num_divorced <- scales::comma(filter(povertytables$marital_ad$rel_numbers, Group == "Divorced / Civil Partnership dissolved / separated")[[params$data_period]])
num_married <- scales::comma(filter(povertytables$marital_ad$rel_numbers, Group == "Married / Civil Partnership")[[params$data_period]])
num_widowed <- scales::comma(filter(povertytables$marital_ad$rel_numbers, Group == "Widowed")[[params$data_period]])
num_cohab <- scales::comma(filter(povertytables$marital_ad$rel_numbers, Group == "Cohabiting")[[params$data_period]])

description <- paste0("In ", params$data_period, ", the relative poverty rate after housing costs was highest for single adults (", rate_single, ", ", num_single, " adults each year) and divorced (or separated) adults (", rate_divorced, ", ", num_divorced, "). Married adults were the least likely to be in poverty (", rate_married, ", ", num_married, "), and widowed and cohabiting adults were in the middle (", rate_widowed, " and ", rate_cohab, "; ", num_widowed, " and ", num_cohab, ").

Poverty among widowed and divorced/separated adults largely decreased over the long term, whereas the trend for singles, cohabiting and married adults was broadly flat over time.

", infobox("By 'Single' we mean adults who have never been married or in a Civil Partnership, and are not living with their partner. The 'Married' category includes Civil Partnerships, and couples who are married or in a Civil Partnership but temporarily living apart. The 'Divorced' category includes divorced couples, dissolved Civil Partnerships, and couples who are married or in a Civil partnership but are not living together because of estrangement."))

```

```{r marital_chart}

message(keymessage)
htmltools::tags$figure(id = paste0("figure-", fign),
                       htmltools::tags$div(class = "highcharts-description",
                                           id = paste0("figure-", fign, "-description"),
                                           create_html(description)),
                       chart)
tables_panel(list(table1, table2))

```

### Ethnicity

Ethnicity data relates to all people in a household and is based on the ethnicity of the adult with the highest income.

```{r ethnic}

keymessage <- "Relative poverty rates higher for ethnic minorities"

fign <- fign + 1

title <- paste0("Figure ", fign, ": ", keymessage)
subtitle <- str_c("Proportion of people in relative poverty after housing costs, Scotland ", params$data_period_5)

chart <- povertycharts$pp_ethnic %>% 
  highcharter::hc_title(text = title) %>% 
  highcharter::hc_subtitle(text = subtitle) %>% 
  highcharter::hc_exporting(filename = paste0("fig_", fign)) %>% 
  highcharter::hc_credits(enabled = TRUE, text = source, href = href)

table1 <- datatable(data = htmltables$pp_ethnic,
                            id = paste0("table-", fign, "a"),
                            caption = paste0("Table ", fign, "a: ", subtitle),
                            source = source,
                            options = list(rowHeaders = TRUE))

table2 <- datatable(data = htmltables$pp_ethnic_age,
                            id = paste0("table-", fign, "b"),
                            caption = paste0("Table ", fign, "b: Median age of household head"),
                            source = source,
                            options = list(rowHeaders = TRUE))

table3 <- datatable(data = htmltables$pp_ethnic_samples,
                            id = paste0("table-", fign, "c"),
                            caption = paste0("Table ", fign, "c: Sample size"),
                            source = source,
                            options = list(rowHeaders = TRUE))

rate_asian <- scales::percent(filter(povertytables$ethgrphh_pp$rel_rates, Group == "Asian or Asian British")[[params$data_period_5]]) 
rate_mixed <- scales::percent(filter(povertytables$ethgrphh_pp$rel_rates, Group == "Mixed, Black or Black British, and Other")[[params$data_period_5]])
rate_brits <- scales::percent(filter(povertytables$ethgrphh_pp$rel_rates, Group == "White - British")[[params$data_period_5]])
rate_white <- scales::percent(filter(povertytables$ethgrphh_pp$rel_rates, Group == "White - Other")[[params$data_period_5]])

num_brits <- scales::comma(filter(povertytables$ethgrphh_pp$rel_numbers, Group == "White - British")[[params$data_period_5]])
num_white <- scales::comma(filter(povertytables$ethgrphh_pp$rel_numbers, Group == "White - Other")[[params$data_period_5]])

age_white <- filter(povertytables$median_ages$ethnic, ethgrphh == "White - Other")[["age"]]
age_brits <- filter(povertytables$median_ages$ethnic, ethgrphh == "White - British")[["age"]]
age_asian <- filter(povertytables$median_ages$ethnic, ethgrphh == "Asian or Asian British")[["age"]]
age_mixed <- filter(povertytables$median_ages$ethnic, ethgrphh == "Mixed, Black or Black British, and Other")[["age"]]

description <- paste0("In ", params$data_period_5, ", people from non-white minority ethnic groups were more likely to be in relative poverty after housing costs compared to those from the 'White - British' and 'White - Other' groups.

The poverty rate was ", rate_asian, " for the 'Asian or Asian British' ethnic groups and ", rate_mixed, " for 'Mixed, Black or Black British and Other' ethnic groups (no population estimates available due to the small sample).

The poverty rate amongst the 'White - Other' group was ", rate_white, " (", num_white, " people) and that of the 'White - British' group was ", rate_brits, " (", num_brits, " people).

", infobox(paste0("This analysis doesn't take into account differences in the age profiles of the ethnic groups.\n\nFor the 'White - British' ethnic group the median average age of the highest income earner was ", age_brits, ", compared with a median age of ", age_white, " for the 'White - Other' ethnic group, ", age_asian, " for 'Asian or Asian British' and ", age_mixed, " for 'Mixed, Black, Black British or Other' ethnic groups.\n\nOlder people have a lower poverty rate, so the age profile partly explains the lower poverty rate for the 'White - British' ethnic group. However, the age difference cannot explain the entire gap in poverty rates between ethnic groups.\n\nThe  ", abbr('ONS', 'Office for National Statistics')," [working paper on 'Equality across different areas of life in the UK'](https://www.ons.gov.uk/methodology/methodologicalpublications/generalmethodology/onsworkingpaperseries/equalityacrossdifferentareasoflifeintheuk2011to2020#ethnic-group) has more information on how age may partly explain different outcomes for different ethnic groups.")))

```

```{r ethnic_chart}

message(keymessage)
htmltools::tags$figure(id = paste0("figure-", fign),
                       htmltools::tags$div(class = "highcharts-description",
                                           id = paste0("figure-", fign, "-description"),
                                           create_html(description)),
                       chart)
tables_panel(list(table1, table2))

```

#### More information on ethnicity data

Figure `r fign` shows an ethnicity breakdown based on an average of data from the past five years. This provides a reasonably detailed breakdown, whilst still using relatively recent data. A more [detailed ethnicity breakdown using ten years of data](https://www.gov.scot/publications/additional-poverty-statistics-2018/) was previously published. 

Due to the small sample sizes for some of the ethnic groups, and the fact that ethnic composition of the population is not accounted for in the survey weighting process, estimates fluctuate between years and the measurement uncertainty is fairly large.

### Religion

Data on religion is available for adults only, so this analysis does not include children.

```{r religion}

keymessage <- "Relative poverty rates higher for Muslims"

fign <- fign + 1

title <- paste0("Figure ", fign, ": ", keymessage)
subtitle <- str_c("Proportion of adults in relative poverty after housing costs, Scotland ", 
                  params$data_period_5)

chart <- povertycharts$ad_religion %>% 
  highcharter::hc_title(text = title) %>% 
  highcharter::hc_subtitle(text = subtitle) %>% 
  highcharter::hc_exporting(filename = paste0("fig_", fign)) %>% 
  highcharter::hc_credits(enabled = TRUE, text = source, href = href)

table1 <- datatable(data = htmltables$ad_religion,
                            id = paste0("table-", fign, "a"),
                            caption = paste0("Table ", fign, "a: ", subtitle),
                            source = source,
                            options = list(rowHeaders = TRUE))

table2 <- datatable(data = htmltables$ad_religion_age,
                            id = paste0("table-", fign, "b"),
                            caption = paste0("Table ", fign, "b: Median age"),
                            source = source,
                            options = list(rowHeaders = TRUE))

table3 <- datatable(data = htmltables$ad_religion_samples,
                            id = paste0("table-", fign, "c"),
                            caption = paste0("Table ", fign, "c: Sample size"),
                            source = source,
                            options = list(rowHeaders = TRUE))

rate_all <- fmtpct(filter(povertytables$religsc_ad$rel_rates, Group == "All")[[params$data_period_5]])

rate_muslim <- scales::percent(filter(povertytables$religsc_ad$rel_rates, Group == "Muslim")[[params$data_period_5]])
rate_cos <- scales::percent(filter(povertytables$religsc_ad$rel_rates, Group == "Church of Scotland")[[params$data_period_5]])
rate_cath <- scales::percent(filter(povertytables$religsc_ad$rel_rates, Group == "Roman Catholic")[[params$data_period_5]]) 
rate_other <- scales::percent(filter(povertytables$religsc_ad$rel_rates, Group == "Other Christian")[[params$data_period_5]])

num_muslim <- scales::comma(filter(povertytables$religsc_ad$rel_numbers, Group == "Muslim")[[params$data_period_5]])
num_cos <- scales::comma(filter(povertytables$religsc_ad$rel_numbers, Group == "Church of Scotland")[[params$data_period_5]])
num_cath <- scales::comma(filter(povertytables$religsc_ad$rel_numbers, Group == "Roman Catholic")[[params$data_period_5]])
num_other <- scales::comma(filter(povertytables$religsc_ad$rel_numbers, Group == "Other Christian")[[params$data_period_5]])

age_cos <- filter(povertytables$median_ages$religion, religsc == "Church of Scotland")[["age"]]
age_muslim <- filter(povertytables$median_ages$religion, religsc == "Muslim")[["age"]]
age_none <- filter(povertytables$median_ages$religion, religsc == "No religion")[["age"]]

description <- paste0("In ", params$data_period_5, ", Muslim adults were more likely to be in relative poverty (", rate_muslim, ", ", num_muslim, " each year) than adults overall (", rate_all, "), after housing costs were taken into account.

Of adults belonging to the Church of Scotland, ", rate_cos, " were in relative poverty after housing costs (", num_cos, " adults each year), compared to ", rate_cath, " of Roman Catholic adults (", num_cath, " adults) and adults of other Christian denominations (", rate_other, "; ", num_other, " adults).

", infobox(paste0("This analysis doesn't take into account differences in the age profiles of the religions.\n\nFor adults belonging to the Church of Scotland, the median average age was ", age_cos , ". In contrast, the median age was ", age_muslim, " for Muslim adults, and ", age_none, " for adults belonging to no religion.\n\nOlder adults have a lower poverty rate, so age profile partly explains the lower poverty rate for adults belonging to the Church of Scotland. However, the age difference cannot explain the entire gap in poverty rates between religious groups.")))

```

```{r religion_chart}

message(keymessage)
htmltools::tags$figure(id = paste0("figure-", fign),
                       htmltools::tags$div(class = "highcharts-description",
                                           id = paste0("figure-", fign, "-description"),
                                           create_html(description)),
                       chart)
tables_panel(list(table1, table2))

```

#### More information on religion data

Figure `r fign` shows an average for data from the past five financial years.

Due to the small sample sizes for some of the religious groups, and the fact that religious composition of the population is not accounted for in the survey weighting process, estimates fluctuate between years and the measurement uncertainty is fairly large.

### Disability

```{r disability1}

keymessage <- "Relative poverty rates higher where a household member is disabled"

fign <- fign + 1

title <- paste0("Figure ", fign, ": ", keymessage)
subtitle <- "Proportion of people in relative poverty after housing costs, Scotland"

chart <- povertycharts$pp_dis1 %>% 
  highcharter::hc_title(text = title) %>% 
  highcharter::hc_subtitle(text = subtitle) %>% 
  highcharter::hc_exporting(filename = paste0("fig_", fign)) %>% 
  highcharter::hc_credits(enabled = TRUE, text = source, href = href)

table1 <- datatable(data = htmltables$pp_dis1,
                            id = paste0("table-", fign, "a"),
                            caption = paste0("Table ", fign, "a: ", subtitle),
                            source = source,
                            options = list(rowHeaders = TRUE))

table2 <- datatable(data = htmltables$pp_dis1_samples,
                            id = paste0("table-", fign, "b"),
                            caption = paste0("Table ", fign, "b: Sample size"),
                            source = source,
                            options = list(rowHeaders = TRUE))

rate_dis <- filter(htmltables$pp_dis1, Group == "Someone disabled")[[params$data_period]]
rate_nodis <- filter(htmltables$pp_dis1, Group == "No-one disabled")[[params$data_period]]

num_dis <- scales::comma(filter(povertytables$disab_pp$rel_numbers, Group == "In household with disabled person(s)")[[params$data_period]])
num_nodis <- scales::comma(filter(povertytables$disab_pp$rel_numbers, Group == "In household with no disabled person(s)")[[params$data_period]])

description <- paste0("Poverty rates remain higher for households in which somebody is disabled compared to those where no-one is disabled. The gap between the two groups has remained fairly steady over the last few years. However, the latest estimate shows an increase for people with disabled household members. It is yet unclear whether this is the beginning of a new trend, or if it is a volatile data point. We do know that the increase is related to the increase of economically inactive working-age adults in poverty, see the discussion in the [working-age section](#working-age-relative-poverty).

In ", params$data_period, ", the poverty rate after housing costs for people in households with a disabled person was ", rate_dis, " (", num_dis, " people each year). This compares with ", rate_nodis, " (", num_nodis, " people) in a household without disabled household members.")

```

```{r pp_disability1_chart}

message(keymessage)
htmltools::tags$figure(id = paste0("figure-", fign),
                       htmltools::tags$div(class = "highcharts-description",
                                           id = paste0("figure-", fign, "-description"),
                                           create_html(description)),
                       chart)
tables_panel(list(table1, table2))

```

`r infobox("The way in which information on disabled people is collected changed several times during this timeseries. This causes breaks in the timeseries. Due to these changes, care needs to be taken when considering long-term trends.\n\nSince 2012/13, disabled people are identified as those who report any physical or mental health condition(s) or illness(es) that last or are expected to last 12 months or more, and which limit their ability to carry out day-to-day activities.\n\nMore detail can be found on pages 34-36 in the [2015/16 Households Below Average Incomes technical report](https://www.gov.uk/government/statistics/households-below-average-income-199495-to-201516).")`

#### Additional living costs of disabled people

Some illnesses and disabilities incur additional living costs. The poverty measure does not normally consider this. However, the analysis shown in Figure `r fign + 1` uses an adjusted poverty rate that partly accounts for additional living costs for those households where someone receives disability benefits.

```{r disability2}

keymessage <- "Relative poverty rates, with disability benefits removed from household income, are higher where a household member is disabled"

fign <- fign + 1

title <- paste0("Figure ", fign, ": ", keymessage)
subtitle <- "Proportion of people in relative poverty (disability benefits removed from household income) after housing costs, Scotland"

chart <- povertycharts$pp_dis2 %>% 
  highcharter::hc_title(text = title) %>% 
  highcharter::hc_subtitle(text = subtitle) %>% 
  highcharter::hc_exporting(filename = paste0("fig_", fign)) %>% 
  highcharter::hc_credits(enabled = TRUE, text = source, href = href)

table1 <- datatable(data = htmltables$pp_dis2,
                            id = paste0("table-", fign, "a"),
                            caption = paste0("Table ", fign, "a: ", subtitle),
                            source = source,
                            options = list(rowHeaders = TRUE))

table2 <- datatable(data = htmltables$pp_dis2_samples,
                            id = paste0("table-", fign, "b"),
                            caption = paste0("Table ", fign, "b: Sample size"),
                            source = source,
                            options = list(rowHeaders = TRUE))

rate_dis <- filter(htmltables$pp_dis2, Group == "Someone disabled")[[params$data_period]]
rate_nodis <- filter(htmltables$pp_dis2, Group == "No-one disabled")[[params$data_period]]

num_dis <- scales::comma(filter(povertytables$disab_nobens_pp$rel_numbers, Group == "In household with disabled person(s)")[[params$data_period]])
num_nodis <- scales::comma(filter(povertytables$disab_nobens_pp$rel_numbers, Group == "In household with no disabled person(s)")[[params$data_period]])

description <- paste0("Research shows that additional costs associated with disability vary greatly in level and nature. There is no general agreement on how to measure these costs.

The analysis in Figure ", fign, " excludes Disability Living Allowance, Attendance Allowance and Personal Independence Payments from total household income. These benefits are paid as a contribution towards the additional living costs for disabled people. If this income is excluded from total household income, then we are able to compare households with and without a disabled household member on a more like for like basis.

In ", params$data_period, ", as in previous years, the poverty rate was higher for individuals in households with a disabled person, when disability-related benefits are not included in the household income. After housing costs, the poverty rate was ", rate_dis, " (", num_dis, " people each year) for people living with a disabled household member, and ", rate_nodis, " (", num_nodis, " people) for those without.")

```

```{r pp_disability2_chart}

message(keymessage)
htmltools::tags$figure(id = paste0("figure-", fign),
                       htmltools::tags$div(class = "highcharts-description",
                                           id = paste0("figure-", fign, "-description"),
                                           create_html(description)),
                       chart)
tables_panel(list(table1, table2))

```

`r infobox("Excluding disability benefits from the analysis changes the poverty threshold. This is because some households now have a lower income, which lowers the median and therefore also the poverty threshold. As a consequence, some households without a disabled person that were just below the poverty threshold in the previous analysis will now be above the threshold. As a result, the poverty rate for households without a disabled person is slightly lower.")`
