**A Scottish Government National Statistics publication**

<div class = "pdf">
  <p style = "margin-bottom: 0;">
    <a href="javascript:window.print();">Print page or download as PDF</a>
  </p>
</div>

Published on `r params$pub_day`

This report presents three-year averaged estimates of the proportion of people, children, working-age adults and pensioners in Scotland living in poverty, and other statistics on household income and income inequality. These estimates are used to monitor progress in reducing poverty, child poverty and income inequality.

The latest estimates are from the period between April 2017 and March 2020. This was before the first `r abbr("UK", "United Kingdom")`-wide lockdown due to the coronavirus (COVID-19) pandemic. Therefore, these statistics do not yet tell us anything about the impact of the pandemic on poverty and income inequality.

```{r}

infobox(c("Revision on 14 June 2021:", "* Figures 15-17 now include the previously missing '3 or more children' category.",
        "* Reference tables now include the previously missing table 24 'Relative child poverty by number of children in the household'."))

```

## Key trends

* Relative poverty appears now stable after a recent small rise. Previously, it had been falling slowly since the nineties.
* Income inequality continues to fluctuate.
* Median household incomes continue to rise.

```{r}

rateA <- filter(data$relpov, yearn == max(yearn), weight == "gs_newpp", 
                key == "After housing costs") %>% select(label) %>% pull()
rateB <- filter(data$palma, yearn == max(yearn), key == "Before housing costs") %>%
  mutate(y = fmtpct(round2(y - 1, 2))) %>% select(y) %>% pull()

incB <- filter(data$medians, yearn == max(yearn), key == "Before housing costs") %>%
  select(label) %>% pull()

div(class = "row fluid-row",
     
      div(class = "col-md-4", 
          div(class = "panel panel-default",
              div(class = "panel-heading",
                  div(class = "panel-title",
                     "Relative poverty stable")),
              div(class = "panel-body",
                  a(href = "#relative-poverty",
                    img(src = "img/chart0a.png", 
                        alt = "Relative poverty section",
                        style = "width: 216px")),
                  create_html(paste0(rateA, " of people were living in relative poverty after housing costs in ", params$data_period, ". Overall, the poverty rate appears to have stopped rising."))))),
     
     div(class = "col-md-4", 
          div(class = "panel panel-default",
              div(class = "panel-heading",
                  div(class = "panel-title",
                     "Income inequality continues to fluctuate")),
              div(class = "panel-body",
                  a(href = "#palma",
                    img(src = "img/chart0b.png", 
                        alt = "Income inequality section",
                        style = "width: 216px")),
                  create_html(paste0("The top 10% of the population had ", rateB," more income in ", params$data_period, " than the bottom 40% combined."))))),
     
     div(class = "col-md-4",
         div(class = "panel panel-default",
             div(class = "panel-heading",
                 div(class = "panel-title",
                    "Household incomes continue to rise")),
             div(class = "panel-body",
                 a(href = "#income-trends",
                   img(src = "img/chart0c.png", 
                        alt = "Household income section",
                        style = "width: 216px")),
                 create_html(paste0("Median income before housing costs in Scotland in ", params$data_period, " was ", incB," a week, continuing the rising trend.")))))
     
 )
 
```

Longer-term trends of poverty and income inequality in the `r abbr("UK", "United Kingdom")`, which are similar to those in Scotland, are available from the Department for Work and Pensions' [Households Below Average Income publication](https://www.gov.uk/government/statistics/households-below-average-income-199495-to-201819) on page 12 of the main report.
