## Child poverty

Children are more likely to be in poverty across all measures compared to adults.

The Child Poverty (Scotland) Act 2017 requires us to report every year on four different child poverty measures. The latest single-year estimates can be found in the [Child poverty summary](cpupdate.html). In this report, however, we show three-year averaged estimates only, which are better for seeing trends.

`r infobox("In this publication, 'child' refers to a dependent child. This is explained in the [Definitions section](#dependent-children).")`

### Trends {.tabset}

#### Relative poverty

```{r}
keymessage <- "Child poverty gradually rising"
message(keymessage)

rateA <- filter(data$relpov, yearn == max(yearn), weight == "gs_newch", 
                key == "After housing costs") %>% select(label) %>% pull()
numberA <- filter(data$relpov, yearn == max(yearn), weight == "gs_newch", 
                key == "After housing costs") %>% select(number) %>% 
  mutate(number = comma(round2(number, -4))) %>% pull()
rateB <- filter(data$relpov, yearn == max(yearn), weight == "gs_newch", 
                key == "Before housing costs") %>% select(label) %>% pull()
numberB <- filter(data$relpov, yearn == max(yearn), weight == "gs_newch", 
                key == "Before housing costs") %>% select(number) %>% 
  mutate(number = comma(round2(number, -4))) %>% pull()
```

It is estimated that `r rateA` of children (`r numberA` children each year) were living in relative poverty after housing costs in `r params$data_period`. Before housing costs, it is estimated that `r rateB` of children (`r numberB` children each year) were in relative poverty.

After a long fall between the late nineties and 2010-13, which slowed briefly just before the recession, child poverty rates have been gradually rising again.

```{r chart10}
title <- str_c("Figure 10: ", keymessage)
subtitle <- "Proportion of children in relative poverty, Scotland"
interactive(povertycharts$chart10, title = title, subtitle = subtitle)
```

Having paid work is an effective way out of poverty, and those families where all adults are in full-time work have a low poverty risk. But having a job is not always enough, for example when it does not pay well, or when someone is unable to work enough hours.

```{r}
keymessage <- "Two thirds of children in poverty live in working households"
message(keymessage)

rateA <- filter(data$workpov, yearn == max(yearn), weight == "gs_newch", 
                key == "After housing costs") %>% 
  mutate(label = fmtpct(round2(composition, 2))) %>%
  select(label) %>% pull()
numberA <- filter(data$workpov, yearn == max(yearn), weight == "gs_newch", 
                key == "After housing costs") %>% select(number) %>% 
  mutate(number = comma(round2(number, -4))) %>% pull()

```

While the poverty risk is much lower for children in working households compared to those in non-working households, not all work pays enough to lift the household above the poverty threshold.

It is estimated that in `r params$data_period`, `r rateA` of children in relative poverty after housing costs were living in working households (`r numberA` children each year). 

This data suggests that after a steady and steep increase of in-work poverty since 2007-10, reflecting high employment rates, in-work poverty is now stable at a high level.

`r infobox("The terms ‘working’ and ‘in-work poverty’ here refer to paid employment. They do not include unpaid work such as caring for your children or other family members. In-work poverty refers to the share of children in poverty who live in households where at least one member of the household is in either full or part-time paid work.")`

```{r chart11}
title <- str_c("Figure 11: ", keymessage)
subtitle <- "Children in relative poverty after housing costs by household work status, Scotland, Scotland"
interactive(povertycharts$chart11, height = 4.5, title = title, subtitle = subtitle)
```

Before-housing-costs data looks similar, and the analysis is available in the associated tables.

#### Absolute poverty

```{r}
keymessage <- "Absolute poverty rate for children largely stable"
message(keymessage)

rateA <- filter(data$abspov, yearn == max(yearn), weight == "gs_newch", 
                key == "After housing costs") %>% select(label) %>% pull()
numberA <- filter(data$abspov, yearn == max(yearn), weight == "gs_newch", 
                key == "After housing costs") %>% select(number) %>% 
  mutate(number = comma(round2(number, -4))) %>% pull()
rateB <- filter(data$abspov, yearn == max(yearn), weight == "gs_newch", 
                key == "Before housing costs") %>% select(label) %>% pull()
numberB <- filter(data$abspov, yearn == max(yearn), weight == "gs_newch", 
                key == "Before housing costs") %>% select(number) %>% 
  mutate(number = comma(round2(number, -4))) %>% pull()
```

Absolute child poverty after housing costs affected `r rateA` (`r numberA` children each year). Before housing costs, absolute child poverty was at `r rateB` (`r numberB` children each year).

Absolute child poverty before housing costs steadily decreased between the mid-nineties and just before the recession. The decrease of the after-housing-costs rate then slowed down. The before-housing-costs measure continued to decrease until 2013-16 and has been largely flat since.

```{r chart12}
title <- str_c("Figure 12: ", keymessage)
caption <- "Proportion of children in absolute poverty, Scotland"
interactive(povertycharts$chart12, height = 4.5, title = title, 
            subtitle = subtitle)
```

#### Material deprivation

```{r}
keymessage <- "Children in combined material deprivation and low income broadly stable"
message(keymessage)

rateA <- filter(data$cmd, yearn == max(yearn), weight == "gs_newch", 
                key == "After housing costs") %>% select(label) %>% pull()
numberA <- filter(data$cmd, yearn == max(yearn), weight == "gs_newch", 
                key == "After housing costs") %>% select(number) %>% 
  mutate(number = comma(round2(number, -4))) %>% pull()
rateB <- filter(data$cmd, yearn == max(yearn), weight == "gs_newch", 
                key == "Before housing costs") %>% select(label) %>% pull()
numberB <- filter(data$cmd, yearn == max(yearn), weight == "gs_newch", 
                key == "Before housing costs") %>% select(number) %>% 
  mutate(number = comma(round2(number, -4))) %>% pull()

```

It is estimated that `r rateA` of children (`r numberA` children each year) were living in combined low income and material deprivation after housing costs in `r params$data_period`. Before housing costs, this was `r rateB` of children (`r numberB` children).

Combined low income and child material deprivation is an additional way of measuring living standards. It is about households who cannot afford basic goods and activities that are seen as necessities in society.

More detail on this can be found in the [Definitions section](#child-material-deprivation).

```{r chart13}
title <- str_c("Figure 13: ", keymessage)
subtitle <- "Proportion of children in combined low income and material deprivation, Scotland"
interactive(povertycharts$chart13, title = title, subtitle = subtitle)
```

`r infobox("Material deprivation data has been collected since 2004/05. Due to a change in the methodology, it is not possible to compare the most recent years with years before 2010-13. This is shown as a break in the chart.")`

#### Food security

```{r}
keymessage <- "Children in poverty often live in households that lack food security"
message(keymessage)

rateH_all <- filter(data$foodsec, x == "All children", key == "High") %>%
  mutate(y = fmtpct(round2(y, 2))) %>% select(y) %>% pull()

rateL_all <- filter(data$foodsec, x == "All children", key == "High") %>%
  mutate(y = fmtpct(round2(1 - y, 2))) %>% select(y) %>% pull()

rateH_rel <- filter(data$foodsec, x == "Children in relative poverty", 
                    key == "High") %>%
  mutate(y = fmtpct(round2(y, 2))) %>% select(y) %>% pull()

rateH_abs <- filter(data$foodsec, x == "Children in absolute poverty", 
                    key == "High") %>%
  mutate(y = fmtpct(round2(y, 2))) %>% select(y) %>% pull()

```

Food security is measured at a household level. The person in the household who knows most about buying and preparing food responds to the questions about food security. Note that not everyone in the household may experience this in the same way. For example, a parent may have worried about running out of food or reduced their own meal sizes, but protected their young children from this experience. 

In `r params$data_year`, `r rateH_all` of children lived in households with high food security. This means that `r rateL_all` of children lived in households with marginal, low or very low food security.

Children in poverty were less likely to have high food security: just `r rateH_rel` of those in relative poverty, and `r rateH_abs` of those in absolute poverty lived in high food security households.

`r infobox("Household food security questions were newly added to the Family Resources Survey in 2019/20. They ask about whether people were worried about running out of food, had to reduce meal sizes or skip meals. More information can be found in the [Definitions](#food-security-2) section.")`

```{r chart14}
title <- str_c("Figure 14: ", keymessage)
subtitle <- str_c("Levels of food security of all children and those in poverty after housing costs, Scotland ",
                 params$data_year)
interactive(povertycharts$chart14, title = title, subtitle = subtitle)
```

### Priority groups {.tabset}

Some types of households with children are known to be at a particularly high risk of poverty. These include households with single parents, three or more children, disabled household members, of a minority ethnic background, with a child aged under one, or a mother aged under 25. These groups do not cover everyone at higher risk of poverty, but taken together, they cover the majority of households with children that are in poverty.

Note that households with mothers aged under 25 are not included in this analysis, because there were too few in the sample to provide robust estimates.

`r infobox("Here, the minority ethnic group includes white minorities such as white people who are not British. The [Equality analysis section](#ethnicity) includes more detailed ethnicity categories.")`

#### Relative poverty

Children in the priority groups were more likely to be in relative poverty compared to all children. This was in particular true for those in ethnic minority households, and those with a single parent in the household.

```{r chart15}
keymessage <- "Children in priority groups have a higher risk of being in relative poverty"
title <- str_c("Figure 15: ", keymessage)
subtitle <- paste0("Proportion of children in relative poverty after housing cost, Scotland ", params$data_period)
interactive(povertycharts$chart15, title = title, subtitle = subtitle)
```

#### Absolute poverty

Children in the priority groups were more likely to be in absolute poverty compared to all children. This was in particular true for those with a single parent in the household and for ethnic minority households.

```{r chart16}
keymessage <- "Children in priority groups have a higher risk of being in absolute poverty"
title <- str_c("Figure 16: ", keymessage)
subtitle <- paste0("Proportion of children in absolute poverty after housing costs, Scotland ", params$data_period)
interactive(povertycharts$chart16, title = title, subtitle = subtitle)
```

#### Material deprivation

Children in the priority groups tended to be more likely to be in combined low income and material deprivation compared to all children, except for those with a child under 1. This was in particular true for those with a single parent in the household.

```{r chart17}
keymessage <- "Children in some priority groups have a higher risk of experiencing combined low income and material deprivation"
title <- str_c("Figure 17: ", keymessage)
subtitle <- paste0("Proportion of children in combined low income and material deprivation, Scotland ", params$data_period)
interactive(povertycharts$chart17, title = title, subtitle = subtitle)
```

