---
title: "Accessibility"
lang: "en"
---

<!-- Include skip link and skip target at the top of the document -->

<script>
  $(document).ready(function(){
    $nav = $('#nav');
    $nav.prepend('<a class=\"skiplink\" href=\"#maincontent\">Skip to main content</a>')
    $head = $('#header');
    $head.prepend('<div id=\"maincontent\" class=\"skiptarget\"></div>')
    });
</script>

```{r setup, include = FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)

knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

```

```{r}
source("R/00_functions_rmd.R")
```

This website is run by the Scottish Government. We want it to be accessible and usable for as many people as possible.

## Feedback

[Contact us](#contact) if you:

- Cannot access any part of this site or want to report an accessibility problem.
- Need information on this website in a different format. Such as accessible PDF, large print, easy read, audio recording or braille. We’ll consider your request and get back to you as soon as we can.

### Enforcement

If you’re not happy with how we respond to your feedback, contact the [Equality Advisory and Support Service](https://www.equalityadvisoryservice.com/). They are an independent advice service. They will advise you on what to do next.

## Compliance

Scottish Government is committed to making its websites accessible, in accordance with the [Public Sector Bodies (Websites and Mobile Applications) (No. 2) Accessibility Regulations 2018](https://www.legislation.gov.uk/uksi/2018/952/regulation/4/made).

This accessibility statement applies to the [data.gov.scot/poverty](https://data.gov.scot/poverty) webpages.

### Compliance status

We believe that this website is compliant with the regulations. This is based on a non-expert assessment using [`r abbr("WCAG", "Web Content Accessibility Guidelines")` 2.1 success criteria and techniques](https://www.w3.org/WAI/WCAG21/quickref/).

### Potentially non-accessible content

The content listed below is compliant with the regulations, and yet may not be fully accessible:

- Some links to external research papers lead to pdf documents which may not be fully accessible, and no other format is available.
- Data for download is available in xlsx format only, but can on request be made available in other formats.
- Most charts are vector graphics which may be difficult to explore using screen readers and other assistive tools.

## Preparation of statement

This statement was prepared on 18 March 2021.

## Contact {#contact}

For enquiries about these reports please contact us.

### Email

social-justice-analysis@gov.scot

### Address

The Scottish Government<br>
Communities Analysis Division<br>
Area 2H North<br>
Victoria Quay<br>
Edinburgh EH6 6QQ

### Phone

0131 244 5851<br>
