---
title: "Child poverty summary"
lang: "en"
description: "Update of Scotland's child poverty target measures"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)

knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

```

```{r}

library(tidyverse)
library(scales)
library(ggiraph)
library(htmltools)

source("R/00_functions.R")
source("R/00_colours.R")

cp_charts <- readRDS("data/cpcharts.rds")

# Theme ----

mytheme <- theme_grey() +
  theme(text = element_text(colour = SGgreys[1], size = 20),

        line = element_line(colour = SGgreys[1],
                            linetype = 1,
                            lineend = 2,
                            size = 0.5),

        plot.caption = element_text(hjust = 1,
                                    margin = margin(t = 20)),

        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),

        axis.line.x = element_line(),
        axis.ticks.length = unit(2, "pt"),
        axis.ticks.y = element_blank(),

        axis.title = element_blank(),
        axis.text.y = element_blank())

theme_set(mytheme)

```

<button onclick = "window.print();" class = "btn noprint" >
  Print or save as PDF
</button>

**Published on 31 March 2022. Next update: March 2023**

The Child Poverty Act 2017 contains four income-based targets to reduce child poverty in Scotland by 2030. The charts show the interim and final targets and the latest poverty estimates and trends. Note that the dots and labels show single-year estimates and the lines show the three-year averages. Single-year estimates give the latest best estimates, and three-year averages show trends more accurately.

`r infobox("Due to data quality issues, only the persistent child poverty estimate was updated in 2022. We expect that all measures will be updated again in 2023.")`

The charts show that relative and absolute child poverty fell between 1994/95, when data collection began, and 2011/12. Since then, relative child poverty has been gradually increasing while absolute child poverty remained largely stable. Child material deprivation has been measured since 2004 and persistent poverty since 2010, neither with an obvious long-term trend. In the most recent period, persistent child poverty appears to have dropped slightly.

```{r panels}

make4panels(title_tl = "Relative poverty",
            subtitle_tl = "= low income relative to the rest of society", 
            text_tl = "Proportion of children living in households with equivalised incomes below 60% of the median (middle) UK income in the current year.", 
            chart_tl = cp_charts$rel, 
            desc_tl = "Line chart of the relative child poverty trend, also including single-year data points and the interim and final targets, showing a small increase in recent years, moving further away from the targets.",
            
            title_tr = "Absolute poverty", 
            subtitle_tr = "= low living standards relative to 2010/11", 
            text_tr = "Proportion of children living in households with equivalised incomes below 60% of the median UK income in 2010/11 adjusted for inflation.", 
            chart_tr = cp_charts$abs, 
            desc_tr = "Line chart of the absolute child poverty trend, also including single-year data points and the interim and final targets, showing an initial fall followed by stagnation in recent years at a level considerably higher than the targets.",
            
            title_bl = "Combined low income and material deprivation", 
            subtitle_bl = "= unable to afford basic necessities", 
            text_bl = "Proportion of children in households with equivalised incomes below 70% of the median UK income and going without certain basic goods and services.", 
            chart_bl = cp_charts$cmd, 
            desc_bl = "Line chart of the proportion of children in combined low income and material deprivation trend, also including single-year data points and the interim and final targets, showing little change and a level that is considerably higher than the targets.",
            
            title_br = "Persistent poverty", 
            subtitle_br = "= in poverty for several years", 
            text_br = "Proportion of children who have lived in relative poverty in three or more of the last four years.", 
            chart_br = cp_charts$per, 
            desc_br = "Line chart of the persistent child poverty trend, also including single-year data points and the interim and final targets with data only available since 2010, showing a broadly stable level over time and a small drop in the most recent period.")

```


[More information and further child poverty analysis](https://www.gov.scot/collections/child-poverty-statistics/) is available.
