---
title: "Child poverty summary"
lang: "en"
description: "Update of Scotland's child poverty target measures"
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(
echo = FALSE,
message = FALSE,
warning = FALSE
)

knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

```

<button onclick = "window.print();" class = "btn noprint" >
Print or save as PDF
</button>

**Published on 23 March 2023. Next update: March 2024**

The Child Poverty Act 2017 contains four income-based targets to reduce child poverty in Scotland by 2030. The charts show the interim and final targets and the latest poverty estimates and trends. The dots and labels show single-year estimates, the shaded areas show indicative 95% confidence intervals around the single-year estimates, and the lines show the three-year averages. Single-year estimates give the latest best estimates, and three-year averages show trends more accurately.

```{r panels}

library(highcharter)
library(htmltools)
source("R/00_functions.R")

# overall description

desc_overall <- "The charts show that relative and absolute child poverty fell between 1994/95, when data collection began, and 2011/12. Since then, relative child poverty gradually increased a little for a few years, but in recent years, both relative and absolute child poverty have shown little change at a level higher than the targets.

Child material deprivation has been measured since 2004 and persistent poverty since 2010. While material deprivation has changed very little at a level slightly higher than the interim target, persistent child poverty has gradually increased.

Note that the latest child material deprivation estimate cannot be directly compared to previous estimates. This is because the latest period covers a time when families were less able to undertake certain activities due to the pandemic, and not necessarily because they couldn't afford to. This changed how people responded to the material deprivation questions."

# Individual chart descriptions

desc_rel <- "Line chart of the relative child poverty trend also including single-year data points and the interim and final targets, showing little change in recent years, at a level higher than the targets."

desc_abs <- "Line chart of the absolute child poverty trend, also including single-year data points and the interim and final targets, showing an initial fall followed by little change in recent years at a level higher than the targets."

desc_cmd <- "Line chart of the proportion of children in combined low income and material deprivation trend, also including single-year data points and the interim and final targets, showing very little change at a level slightly higher than the interim target."

desc_per <- "Line chart of the persistent child poverty trend, also including single-year data points and the interim and final targets with data only available since 2010, showing a gradual increase over time, moving further away from the targets."

# Definitions

text_rel <- "Definition: Proportion of children living in households with equivalised incomes below 60% of the median (middle) UK income in the current year."
text_abs <- "Definition: Proportion of children living in households with equivalised incomes below 60% of the median UK income in 2010/11 adjusted for inflation."
text_cmd <- "Definition: Proportion of children in households with equivalised incomes below 70% of the median UK income and going without certain basic goods and services."
text_per <- "Definition: Proportion of children who have lived in relative poverty in three or more of the last four years."

cp_charts <- readRDS("data/cpcharts.rds")

tags$figure(role = "figure",
            div(class = "highcharts-description",
                create_html(desc_overall)),
            
            div(class = "row fluid-row",
                div(class = "col-md-6",
                    
                    div(class = "panel panel-default",
                        div(class = "panel-body",
                            style = "max-width: 438px;",
                            "aria-label" = desc_rel,
                            cp_charts$rel,
                            tags$p(text_rel,
                                   style = "font-size: larger; line-height: inherit;"))),
                    
                    div(class = "panel panel-default",
                        div(class = "panel-body",
                            style = "max-width: 438px;",
                            "aria-label" = desc_cmd,
                            cp_charts$cmd,
                            tags$p(text_cmd,
                                   style = "font-size: larger; line-height: inherit;")))
                ),
                
                div(class = "col-md-6",
                    
                    div(class = "panel panel-default",
                        div(class = "panel-body",
                            style = "max-width: 438px;",
                            "aria-label" = desc_abs,
                            cp_charts$abs,
                            tags$p(text_abs,
                                   style = "font-size: larger; line-height: inherit;"))),
                    
                    div(class = "panel panel-default",
                        div(class = "panel-body",
                            style = "max-width: 438px;",
                            "aria-label" = desc_per,
                            cp_charts$per,
                            tags$p(text_per,
                                   style = "font-size: larger; line-height: inherit;")))
                )
            )
)


```

More information and further child poverty analysis is available in the [main report](index.html#children) and on the Scottish Government [child poverty analysis pages](https://www.gov.scot/collections/child-poverty-statistics/).
