---
title: "Download data"
lang: "en"
description: "Data download for the official poverty and income inequality statistics in Scotland"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)

knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r}
source("R/00_functions.R")
```

## Poverty and child poverty data

The data used to produce the charts of the main report is available for download below. The files also contain further poverty and child poverty analysis which is not shown in the charts of the main report.

### Three-year averages

All estimates used in the main report except for the data in Figure 36, plus additional poverty and child poverty breakdowns, are available for download. Estimates are averages of three years' worth of data in order to reduce data volatility and show trends more accurately.

<button onclick="location.href='xls/data2023.xlsx'" type="button" class = "btn noprint">
  Download, XLSX format (1 MB)
</button>
<button onclick="location.href='xls/data2023.ods'" type="button" class = "btn noprint">
  Download, ODS format (330 KB)
</button>

### Single-year estimates

In addition, estimates of the headline indicators and household income based on single years of data are also available. These are used in Figures 34 and 35 of the main report, and in the [child poverty update](cpupdate.html). They are not used in any other charts.

<button onclick="location.href='xls/data2023_1yr.xlsx'" type="button" class = "btn noprint">
  Download, XLSX format (260 KB)
</button>
<button onclick="location.href='xls/data2023_1yr.ods'" type="button" class = "btn noprint">
  Download, ODS format (60 KB)
</button>

### Measurement uncertainty

Measurement uncertainty information for the headline estimates is available.

- [confidence intervals](uncertainty.html)

## Persistent poverty data

The data used to produce the charts of the Persistent Poverty in Scotland report is available for download.

<button onclick="location.href='xls/data2023_persistent.xlsx'" type="button" class = "btn noprint">
  Download, XLSX format (10 KB)
</button>
<button onclick="location.href='xls/data2023_persistent.ods'" type="button" class = "btn noprint">
  Download, ODS format (50 KB)
</button>


