## Persistent poverty

Someone is in persistent poverty if they have been in poverty for three or more of the last four years. This measure is important because the longer someone is in poverty, the more it impacts on their health, well-being, and overall life chances.

Poverty estimates in this publication are derived by looking at household income before housing costs are paid for (BHC) and after housing costs are paid for (AHC). In Scotland, after-housing-costs poverty indicators are more commonly used.

### All individuals {.tabset}

- Between `r params$data_period_long`, `r persistenttables[["tableScotland"]]$pp %>% tail(1L)` of people in Scotland were in persistent poverty after housing costs, the same as in the previous period (`r params$prev_period`). 
- Scotland had levels of persistent poverty after housing costs that were similar to Wales (both 12%) and England (13%), and slightly higher than Northern Ireland (9%). 

#### After housing costs

```{r chart01}

title <- "Figure 1: Proportion of people in persistent poverty after housing costs by UK country"
subtitle = NULL
interactive(persistentcharts[["chart01"]], title = title, subtitle = subtitle,
            height = 4)
```

#### Before housing costs

```{r chart02}
title <- "Figure 2: Proportion of people in persistent poverty before housing costs by UK country"
subtitle = NULL
interactive(persistentcharts[["chart02"]], title = title, subtitle = subtitle,
            height = 4)
```

#### Tables

```{r table1}

caption <- "Proportion of people in persistent poverty after housing costs"
data <- persistenttables[["table1"]]
pers_source <- persistenttables[["source"]]

kableExtra::kable(data, caption = caption, align = c('lrrrrr')) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"),
                full_width = F, position = "left")  %>%
  footnote(general = pers_source,
           general_title = "") %>%
  add_a11y()

```

```{r table2}

caption <- "Proportion of people in persistent poverty before housing costs"
data <- persistenttables[["table2"]]

kableExtra::kable(data, caption = caption, align = c('lrrrrr')) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"),
                full_width = F, position = "left")  %>%
  footnote(general = pers_source,
           general_title = "") %>%
  add_a11y()

```

### Children {.tabset}

- `r persistenttables[["tableScotland"]]$ch %>% tail(1L)` of children in Scotland were in persistent poverty after housing costs in `r params$data_period`, compared to `r persistenttables[["tableScotland"]]$ch %>% tail(2L) %>% head(1L)` in the previous period.  
- Children have consistently had the highest risk of living in persistent poverty when comparing with working-age adults and pensioners, both before and after housing costs.
- The persistent child poverty rate after housing costs in Scotland (16% in `r params$data_period`) and Wales (15%) was slightly higher than in Northern Ireland (12%), and slightly lower than in England (19%).

#### After housing costs

```{r chart03}
title <- "Figure 3: Proportion of children in persistent poverty after housing costs by UK country"
subtitle = NULL
interactive(persistentcharts[["chart03"]], title = title, subtitle = subtitle,
            height = 4)
```

#### Before housing costs

```{r chart04}
title <- "Figure 4: Proportion of children in persistent poverty before housing costs by UK country"
subtitle = NULL
interactive(persistentcharts[["chart04"]], title = title, subtitle = subtitle,
            height = 4)
```

#### Tables

```{r table3}

caption <- "Proportion of children in persistent poverty after housing costs"
data <- persistenttables[["table3"]]

kableExtra::kable(data, caption = caption, align = c('lrrrrr')) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"),
                full_width = F, position = "left")  %>%
  footnote(general = pers_source,
           general_title = "") %>%
  add_a11y()

```

```{r table4}

caption <- "Proportion of children in persistent poverty before housing costs"
data <- persistenttables[["table4"]]

kableExtra::kable(data, caption = caption, align = c('lrrrrr')) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"),
                full_width = F, position = "left")  %>%
  footnote(general = pers_source,
           general_title = "") %>%
  add_a11y()

```

### Working-age adults {.tabset}

- Between `r params$data_period_long`, `r persistenttables[["tableScotland"]]$wa %>% tail(1L)` of working-age adults in Scotland were in persistent poverty after housing costs, unchanged from the previous period.
- Persistent poverty rates for working-age adults were similar (between 9% and 12%) in all countries.

#### After housing costs

```{r chart05}
title <- "Figure 5: Proportion of working-age adults in persistent poverty after housing costs by UK country"
subtitle = NULL
interactive(persistentcharts[["chart05"]], title = title, subtitle = subtitle,
            height = 4)
```

#### Before housing costs

```{r chart06}
title <- "Figure 6: Proportion of working-age adults in persistent poverty before housing costs by UK country"
subtitle = NULL
interactive(persistentcharts[["chart06"]], title = title, subtitle = subtitle,
            height = 4)
```

#### Tables

```{r table5}

caption <- "Proportion of working-age adults in persistent poverty after housing costs"
data <- persistenttables[["table5"]]

kableExtra::kable(data, caption = caption, align = c('lrrrrr')) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"),
                full_width = F, position = "left")  %>%
  footnote(general = pers_source,
           general_title = "") %>%
  add_a11y()

```

```{r table6}

caption <- "Proportion of working-age adults in persistent poverty before housing costs"
data <- persistenttables[["table6"]]

kableExtra::kable(data, caption = caption, align = c('lrrrrr')) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"),
                full_width = F, position = "left")  %>%
  footnote(general = pers_source,
           general_title = "") %>%
  add_a11y()

```

### Pensioners {.tabset}

- Between `r params$data_period_long`, `r persistenttables[["tableScotland"]]$pn %>% tail(1L)` of pensioners in Scotland were in persistent poverty after housing costs, unchanged from the previous period.
- For most groups of the population, the persistent poverty rate after housing costs is greater or the same than that before housing costs. For pensioners, however, the opposite is often true, or the rates are very similar. The majority of pensioners own their own home and so have lower housing costs.
- Between `r params$data_period_long`, persistent pensioner poverty levels after housing costs in Scotland (12%) were similar compared to England and Wales (both 11%) and higher than in Northern Ireland (6%).

#### After housing costs

```{r chart07}
title <- "Figure 7: Proportion of pensioners in persistent poverty after housing costs by UK country"
subtitle = NULL
interactive(persistentcharts[["chart07"]], title = title, subtitle = subtitle,
            height = 4)
```

#### Before housing costs

```{r chart08}
title <- "Figure 8: Proportion of pensioners in persistent poverty before housing costs by UK country"
subtitle = NULL
interactive(persistentcharts[["chart08"]], title = title, subtitle = subtitle,
            height = 4)
```

#### Tables

```{r table7}

caption <- "Proportion of pensioners in persistent poverty after housing costs"
data <- persistenttables[["table7"]]

kableExtra::kable(data, caption = caption, align = c('lrrrrr')) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"),
                full_width = F, position = "left")  %>%
  footnote(general = pers_source,
           general_title = "") %>%
  add_a11y()

```

```{r table8}

caption <- "Proportion of pensioners in persistent poverty before housing costs"
data <- persistenttables[["table8"]]

kableExtra::kable(data, caption = caption, align = c('lrrrrr')) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"),
                full_width = F, position = "left")  %>%
  footnote(general = pers_source,
           general_title = "") %>%
  add_a11y()

```
