## Persistent poverty

Someone is in persistent poverty if they have been in poverty for three or more of the last four years. This measure is important because the longer someone is in poverty, the more it impacts on their health, well-being, and overall life chances.

Poverty estimates in this publication are derived by looking at household income before housing costs are paid for (BHC) and after housing costs are paid for (AHC). In Scotland, after-housing-costs poverty indicators are more commonly used.

### All individuals {.tabset}

```{r pp}

fign <- 1
title <- "Proportion of people in persistent poverty, Scotland"
source <- "Source: Understanding Society Survey"
href <- "https://data.gov.scot/poverty/2022/index.html#Data_source"

chart <- persistentcharts$chart1 %>% 
  highcharter::hc_title(text = paste0("Figure ", fign, ": ", title)) %>% 
  highcharter::hc_exporting(filename = paste0("fig_", fign)) %>% 
  highcharter::hc_credits(enabled = TRUE, text = source, href = href)

table <- datatable(data = persistenttables$table1,
                            id = paste0("table-", fign),
                            caption = paste0("Table ", fign, ": ", title),
                            source = source,
                            options = list(rowHeaders = TRUE))

description <- paste0("Between ", params$data_period_long, ", ", persistenttables$tableScotland$pp %>% tail(1L), " of people were in persistent poverty after housing costs, unchanged from the previous period (", params$prev_period, "). ")

```

```{r pp_chart}

htmltools::tags$figure(id = paste0("figure-", fign),
                       htmltools::tags$div(class = "highcharts-description",
                                           id = paste0("figure-", fign, "-description"),
                                           create_html(description)),
                       chart)
tables_panel(table)

```


### Children {.tabset}

```{r ch}

fign <- 1 + fign

title <- "Proportion of children in persistent poverty, Scotland"

chart <- persistentcharts$chart2 %>% 
  highcharter::hc_title(text = paste0("Figure ", fign, ": ", title)) %>% 
  highcharter::hc_exporting(filename = paste0("fig_", fign)) %>% 
  highcharter::hc_credits(enabled = TRUE, text = source, href = href)

table <- datatable(data = persistenttables$table2,
                            id = paste0("table-", fign),
                            caption = paste0("Table ", fign, ": ", title),
                            source = source,
                            options = list(rowHeaders = TRUE))

description <- paste0(persistenttables$tableScotland$ch %>% tail(1L), " of children were in persistent poverty after housing costs in ", params$data_period, ", compared to ", persistenttables$tableScotland$ch %>% tail(2L) %>% head(1L), " in the previous period.

", infobox("Note that previously published estimates have been revised upwards as more accurate data has become available. The Understanding Society dataset gets revised routinely when households re-enter the longitudinal sample and data gaps can be filled. This can affect persistent poverty estimates. Furthermore, an addition to the sample size since 2015 has led to further improvements to existing estimates. Both of these factors are likely to have contributed to the more recent revisions. More information is available in the [Revisions section](#revisions).")
)

```

```{r ch_chart}

htmltools::tags$figure(id = paste0("figure-", fign),
                       htmltools::tags$div(class = "highcharts-description",
                                           id = paste0("figure-", fign, "-description"),
                                           create_html(description)),
                       chart)
tables_panel(table)

```


### Working-age adults {.tabset}

```{r wa}
fign <- 1 + fign

title <- "Proportion of working-age adults in persistent poverty, Scotland"

chart <- persistentcharts$chart3 %>% 
  highcharter::hc_title(text = paste0("Figure ", fign, ": ", title)) %>% 
  highcharter::hc_exporting(filename = paste0("fig_", fign)) %>% 
  highcharter::hc_credits(enabled = TRUE, text = source, href = href)

table <- datatable(data = persistenttables$table3,
                            id = paste0("table-", fign),
                            caption = paste0("Table ", fign, ": ", title),
                            source = source,
                            options = list(rowHeaders = TRUE))

description <- paste0("Between ", params$data_period_long, ", ", persistenttables$tableScotland$wa %>% tail(1L), " of working-age adults were in persistent poverty after housing costs, similar to the previous period.")

```

```{r wa_chart}

htmltools::tags$figure(id = paste0("figure-", fign),
                       htmltools::tags$div(class = "highcharts-description",
                                           id = paste0("figure-", fign, "-description"),
                                           create_html(description)),
                       chart)

tables_panel(table)
```

### Pensioners {.tabset}

```{r pn}
fign <- 1 + fign

title <- "Proportion of pensioners in persistent poverty, Scotland"

chart <- persistentcharts$chart4 %>% 
  highcharter::hc_title(text = paste0("Figure ", fign, ": ", title)) %>% 
  highcharter::hc_exporting(filename = paste0("fig_", fign)) %>% 
  highcharter::hc_credits(enabled = TRUE, text = source, href = href)

table <- datatable(data = persistenttables$table4,
                            id = paste0("table-", fign),
                            caption = paste0("Table ", fign, ": ", title),
                            source = source,
                            options = list(rowHeaders = TRUE))

description <- paste0("Between ", params$data_period_long, ", ", persistenttables$tableScotland$pn %>% tail(1L), " of pensioners were in persistent poverty after housing costs, unchanged from the previous period.")
```

```{r pn_chart}

htmltools::tags$figure(id = paste0("figure-", fign),
                       htmltools::tags$div(class = "highcharts-description",
                                           id = paste0("figure-", fign, "-description"),
                                           create_html(description)),
                       chart)
tables_panel(table)

```


