# In the UK

`r devnote("update manually and make sure that wording makes sense")`

## All individuals {.tabset}

When looking at the overall population, Scotland had levels of persistent poverty after housing costs that were similar to England and Wales and slightly higher than in Northern Ireland.

Between `r params$data_period_long`, Scotland, England and Wales all had a persistent poverty rate after housing cost of 13%, whereas Northern Ireland had a slightly lower rate with 11%. 

### After housing costs

#### Chart 3: Proportion of people in persistent poverty by UK country
```{r chart03}
interactive(persistentcharts[["chart03"]])
```

### Before housing costs

#### Chart 4: Proportion of people in persistent poverty by UK country
```{r chart04}
interactive(persistentcharts[["chart04"]])
```

### Table

#### Table 2: Proportion of people in persistent poverty by country
```{r table2}

data <- persistenttables[["table2"]]
pers_source <- persistenttables[["source"]]

kableExtra::kable(data, 
      col.names = c("Period", "Scotland", "England", "Wales", 
                    "Northern Ireland", "UK", "Scotland", "England", "Wales", 
                    "Northern Ireland", "UK"),
      align = c('lrrrrrrrrrr')) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive", "condensed"),
                full_width = F, position = "left")  %>%
  add_header_above(c(" ", "Before housing costs" = 5, "After housing costs" = 5)) %>%
  footnote(general = pers_source,
           general_title = "")
```

## Children {.tabset}

Between `r params$data_period_long`, the persistent child poverty rate after housing costs in Scotland and Northern Ireland (both 17%) was lower than in England (20%) and Wales (19%). 

In previous periods, persistent child poverty levels in Scotland and Northern Ireland had been similar, and consistently below the English and Welsh rates. 

### After housing costs

#### Chart 5: Proportion of children in persistent poverty by UK country
```{r chart05}
interactive(persistentcharts[["chart05"]])
```

### Before housing costs

#### Chart 6: Proportion of children in persistent poverty by UK country
```{r chart06}
interactive(persistentcharts[["chart06"]])
```

### Table

#### Table 3: Proportion of children in persistent poverty by country
```{r table3}

data <- persistenttables[["table3"]]
pers_source <- persistenttables[["source"]]

kableExtra::kable(data, 
      col.names = c("Period", "Scotland", "England", "Wales", 
                    "Northern Ireland", "UK", "Scotland", "England", "Wales", 
                    "Northern Ireland", "UK"),
      align = c('lrrrrrrrrrr')) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive", "condensed"),
                full_width = F, position = "left")  %>%
  add_header_above(c(" ", "Before housing costs" = 5, "After housing costs" = 5)) %>%
  footnote(general = pers_source,
           general_title = "")
```

## Working-age adults {.tabset}

Between `r params$data_period_long`, persistent poverty rates after housing costs for working-age adults in Scotland were the same as those in Northern Ireland (11%) and slightly below those in England and Wales (both 12%). 

### After housing costs

#### Chart 7: Proportion of working-age adults in persistent poverty by UK country
```{r chart07}
interactive(persistentcharts[["chart07"]])
```

### Before housing costs

#### Chart 8: Proportion of working-age adults in persistent poverty by UK country
```{r chart08}
interactive(persistentcharts[["chart08"]])
```

### Table

#### Table 4: Proportion of working-age adults in persistent poverty by country
```{r table4}

data <- persistenttables[["table4"]]
pers_source <- persistenttables[["source"]]

kableExtra::kable(data, 
      col.names = c("Period", "Scotland", "England", "Wales", 
                    "Northern Ireland", "UK", "Scotland", "England", "Wales", 
                    "Northern Ireland", "UK"),
      align = c('lrrrrrrrrrr')) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive", "condensed"),
                full_width = F, position = "left")  %>%
  add_header_above(c(" ", "Before housing costs" = 5, "After housing costs" = 5)) %>%
  footnote(general = pers_source,
           general_title = "")
```

## Pensioners {.tabset}

Between `r params$data_period_long`, persistent pensioner poverty after housing costs in Scotland (12%) was slightly higher compared to England and Wales (both 11%) and considerably higher compared to Northern Ireland (5%).

### After housing costs

#### Chart 9: Proportion of pensioners in persistent poverty by UK country
```{r chart09}
interactive(persistentcharts[["chart09"]])
```

### Before housing costs

#### Chart 10: Proportion of pensioners in persistent poverty by UK country
```{r chart10}
interactive(persistentcharts[["chart10"]])
```

### Table

#### Table 5: Proportion of pensioners in persistent poverty by country
```{r table5}

data <- persistenttables[["table5"]]
pers_source <- persistenttables[["source"]]

kableExtra::kable(data, 
      col.names = c("Period", "Scotland", "England", "Wales", 
                    "Northern Ireland", "UK", "Scotland", "England", "Wales", 
                    "Northern Ireland", "UK"),
      align = c('lrrrrrrrrrr')) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive", "condensed"),
                full_width = F, position = "left")  %>%
  add_header_above(c(" ", "Before housing costs" = 5, "After housing costs" = 5)) %>%
  footnote(general = pers_source,
           general_title = "")
```
