## Persistent poverty

Someone is in persistent poverty if they have been in poverty for three or more of the last four years. This measure is important because the longer someone is in poverty, the more it impacts on their health, well-being, and overall life chances.

Poverty estimates in this publication are derived by looking at household income before housing costs are paid for (BHC) and after housing costs are paid for (AHC). In Scotland, after-housing-costs poverty indicators are more commonly used.

### All individuals {.tabset}

- Between `r params$data_period_long`, `r persistenttables$tableScotland$pp %>% tail(1L)` of people were in persistent poverty after housing costs, broadly similar to the previous period (`r params$prev_period`). 

```{r chart1}

fign <- 1
title <- "Proportion of people in persistent poverty, Scotland"

interactive(persistentcharts$chart1, 
            title = paste0("Figure ", fign, ": ", title), 
            subtitle = NULL)

table <- datatable(data = persistenttables$table1,
                            id = paste0("table-", fign),
                            caption = paste0("Table ", fign, ": ", title),
                            source = persistenttables$source,
                            options = list(rowHeaders = TRUE))

tables_panel(table)

```


### Children {.tabset}

- `r persistenttables$tableScotland$ch %>% tail(1L)` of children were in persistent poverty after housing costs in `r params$data_period`, compared to `r persistenttables$tableScotland$ch %>% tail(2L) %>% head(1L)` in the previous period.
- This represents a relatively large drop compared to previous estimates, and not all of this decrease is likely to be real. Persistent poverty estimates do tend to fluctuate. They also get revised when households re-enter the longitudinal sample and data gaps can be filled. However, some of the decrease is plausible in part due to increased financial support during the pandemic for some lower-income households. At the same time, reduced earnings and job losses may have resulted in a lower median income. This may have led to a fall in the poverty line, and thereby a drop in the relative poverty rate.

```{r chart2}

fign <- 1 + fign

title <- "Proportion of children in persistent poverty, Scotland"

interactive(persistentcharts$chart2, 
            title = paste0("Figure ", fign, ": ", title), 
            subtitle = NULL)

table <- datatable(data = persistenttables$table2,
                            id = paste0("table-", fign),
                            caption = paste0("Table ", fign, ": ", title),
                            source = persistenttables$source,
                            options = list(rowHeaders = TRUE))

tables_panel(table)

```


### Working-age adults {.tabset}

- Between `r params$data_period_long`, `r persistenttables$tableScotland$wa %>% tail(1L)` of working-age adults were in persistent poverty after housing costs, similar to the previous period.

```{r chart3}
fign <- 1 + fign

title <- "Proportion of working-age adults in persistent poverty, Scotland"

interactive(persistentcharts$chart3, 
            title = paste0("Figure ", fign, ": ", title), 
            subtitle = NULL)

table <- datatable(data = persistenttables$table3,
                            id = paste0("table-", fign),
                            caption = paste0("Table ", fign, ": ", title),
                            source = persistenttables$source,
                            options = list(rowHeaders = TRUE))
tables_panel(table)
```

### Pensioners {.tabset}

- Between `r params$data_period_long`, `r persistenttables$tableScotland$pn %>% tail(1L)` of pensioners were in persistent poverty after housing costs, unchanged from the previous period.

```{r chart4}
fign <- 1 + fign

title <- "Proportion of pensioners in persistent poverty, Scotland"

interactive(persistentcharts$chart4, 
            title = paste0("Figure ", fign, ": ", title), 
            subtitle = NULL)

table <- datatable(data = persistenttables$table4,
                            id = paste0("table-", fign),
                            caption = paste0("Table ", fign, ": ", title),
                            source = persistenttables$source,
                            options = list(rowHeaders = TRUE))
tables_panel(table)
```


