# In the UK

## All individuals {.tabset}

When looking at the overall population, Scotland had levels of persistent poverty after housing costs that were similar to England and Wales and slightly higher than in Northern Ireland.

Between 2014 and 2018, Scotland, England and Wales all had a persistent poverty rate after housing cost of 13%, whereas Northern Ireland had a slightly lower rate with 11%. 

### After housing costs

#### Chart 3: Proportion of people in persistent poverty by UK country
```{r chart03}
interactive(persistentcharts[["chart03"]])
```

### Before housing costs

#### Chart 4: Proportion of people in persistent poverty by UK country
```{r chart04}
interactive(persistentcharts[["chart04"]])
```

### Table

#### Table 2: Proportion of people in persistent poverty by country
```{r table2}
kableExtra::kable(pers_table2, 
      col.names = c("Period", "Scotland", "England", "Wales", 
                    "Northern Ireland", "UK", "Scotland", "England", "Wales", 
                    "Northern Ireland", "UK"),
      align = c('lrrrrrrrrrr')) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive", "condensed"),
                full_width = F, position = "left")  %>%
  add_header_above(c(" ", "Before housing costs" = 5, "After housing costs" = 5)) %>%
  footnote(general = pers_source,
           general_title = "")
```

## Children {.tabset}

Between 2014 and 2018, the persistent child poverty rate after housing costs in Scotland and Northern Ireland (both 17%) was lower than in England (20%) and Wales (19%). 

In previous periods, persistent child poverty levels in Scotland and Northern Ireland had been similar, and consistently below the English and Welsh rates. 

### After housing costs

#### Chart 5: Proportion of children in persistent poverty by UK country
```{r chart05}
interactive(persistentcharts[["chart05"]])
```

### Before housing costs

#### Chart 6: Proportion of children in persistent poverty by UK country
```{r chart06}
interactive(persistentcharts[["chart06"]])
```

### Table

#### Table 3: Proportion of children in persistent poverty by country
```{r table3}
kableExtra::kable(pers_table3, 
      col.names = c("Period", "Scotland", "England", "Wales", 
                    "Northern Ireland", "UK", "Scotland", "England", "Wales", 
                    "Northern Ireland", "UK"),
      align = c('lrrrrrrrrrr')) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive", "condensed"),
                full_width = F, position = "left")  %>%
  add_header_above(c(" ", "Before housing costs" = 5, "After housing costs" = 5)) %>%
  footnote(general = pers_source,
           general_title = "")
```

## Working-age adults {.tabset}

Between 2014 and 2018, persistent poverty rates after housing costs for working-age adults in Scotland were the same as those in Northern Ireland (11%) and slightly below those in England and Wales (both 12%). 

### After housing costs

#### Chart 7: Proportion of working-age adults in persistent poverty by UK country
```{r chart07}
interactive(persistentcharts[["chart07"]])
```

### Before housing costs

#### Chart 8: Proportion of working-age adults in persistent poverty by UK country
```{r chart08}
interactive(persistentcharts[["chart08"]])
```

### Table

#### Table 4: Proportion of working-age adults in persistent poverty by country
```{r table4}
kableExtra::kable(pers_table4, 
      col.names = c("Period", "Scotland", "England", "Wales", 
                    "Northern Ireland", "UK", "Scotland", "England", "Wales", 
                    "Northern Ireland", "UK"),
      align = c('lrrrrrrrrrr')) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive", "condensed"),
                full_width = F, position = "left")  %>%
  add_header_above(c(" ", "Before housing costs" = 5, "After housing costs" = 5)) %>%
  footnote(general = pers_source,
           general_title = "")
```

## Pensioners {.tabset}

Between 2014 and 2018, persistent pensioner poverty after housing costs in Scotland (12%) was slightly higher compared to England and Wales (both 11%) and considerably higher compared to Northern Ireland (5%).

### After housing costs

#### Chart 9: Proportion of pensioners in persistent poverty by UK country
```{r chart09}
interactive(persistentcharts[["chart09"]])
```

### Before housing costs

#### Chart 10: Proportion of pensioners in persistent poverty by UK country
```{r chart10}
interactive(persistentcharts[["chart10"]])
```

### Table

#### Table 5: Proportion of pensioners in persistent poverty by country
```{r table5}
kableExtra::kable(pers_table5, 
      col.names = c("Period", "Scotland", "England", "Wales", 
                    "Northern Ireland", "UK", "Scotland", "England", "Wales", 
                    "Northern Ireland", "UK"),
      align = c('lrrrrrrrrrr')) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive", "condensed"),
                full_width = F, position = "left")  %>%
  add_header_above(c(" ", "Before housing costs" = 5, "After housing costs" = 5)) %>%
  footnote(general = pers_source,
           general_title = "")
```
